KPIs and Visualisations

### 2022-07-24 Report/KPIs and Visualisations.Rmd
Updating with code from the dashboard.

#### Fundraising ROI
```{r}
# Creating Fundraising ROI value
KPI_fundraising_num <- reactive({
  # testing for data quality, returning NA if bad
  if (datatest() == TRUE) {
    NA
  } else {
    exceldata() |>
      # most recent year
      filter(year == max(year)) |>
      transmute(
        # Calculating the value as a percentage; removing all other
        KPI_fundraising_num =
          (revenue_private_contributions +
            revenue_membership_fees - expenses_fundraising) /
            expenses_fundraising * 100
      ) |>
      # just want the single number
      pull() |>
      first()
  }
})
```

#### OPX KPI
```{r}
# Creating OPX KPI
KPI_OPX_num <- reactive({
  # testing for data quality, returning NA if bad
  if (datatest() == TRUE) {
    NA
  } else {
    exceldata() |>
      # most recent year
      filter(year == max(year)) |>
      # Calculating the value as a %
      transmute(
        KPI_OPX = (revenue_private_contributions + revenue_investments +
          assets_cash_in_bank_and_cash_equivalents) /
          (liability_short_term_grants_payable +
            liability_revocable_endowments +
            liability_long_term_grants_payable) * 100
      ) |>
      # Just the single number
      pull() |>
      first()
  }
})
```

@@

#### public/private proportion
```{r}
# Creating public/private proportion
KPI_proportion <- reactive({
  # testing for data quality, returning NA if bad
  if (datatest() == TRUE) {
    NA
  } else {
    exceldata() |>
      transmute(
        proportion = (revenue_private_contributions + revenue_membership_fees) / (revenue_public_grants + revenue_private_contributions + revenue_membership_fees) * 100
      ) |>
      pull() |>
      first()
  }
})
```

@@

# Function to read Excel file
widen_excel <- function(file) {
  read_excel(file) |>
    unite( # uniting Category and Name to deal with multiple 'Other'
      entry, c(Category, Name),
      sep = "_"
    ) |>
    select(-`Sub-category`, -Firm) |> # deleting Sub-category and Firm
    # pivoting
    pivot_wider(names_from = entry, values_from = Value) |>
    clean_names() # cleaning names
}

@@

##### no_information
```{r}
# Function to print a 'no information' plot, if necessary
no_information <- function() {
  ggplot() +
    # text geom with a space to create the plot, but otherwise show nothing
    geom_text(aes(0, 0, label = " ")) +
    # as blank a slate as possible
    theme_void() +
    annotate(
      "text",
      label = "No information available",
      # Using the same palette as the other non-profit/loss plots
      x = 0, y = 0, size = 8, colour = met.brewer("Hokusai2", 1)
    )
}
```

##### bad_data
```{r}
# Function to print a 'bad data' plot, if necessary
bad_data <- function() {
  ggplot() +
    # text geom with a space to create the plot, but otherwise show nothing
    geom_text(aes(0, 0, label = " ")) +
    # as blank a slate as possible
    theme_void() +
    annotate(
      "text",
      label = "Something is wrong\nwith the Excel data.",
      # Using the same palette as the other non-profit/loss plots
      x = 0, y = 0, size = 8, colour = met.brewer("Hokusai2", 1)
    )
}
```

>>>>

#### Fundraising ROI
```{r}
# Creating Fundraising ROI value
KPI_fundraising_num <- reactive({
  # testing for data quality, returning NA if bad
  if (datatest() == TRUE) {
    NA
  } else {
    exceldata() |>
      # most recent year
      filter(year == max(year)) |>
      transmute(
        # Calculating the value as a percentage; removing all other
        KPI_fundraising_num =
          (sum(revenue_private_contributions,
            revenue_membership_fees,
            na.rm = TRUE
          ) - expenses_fundraising) /
            expenses_fundraising * 100
      ) |>
      # just want the single number
      pull() |>
      first()
  }
})
```

#### OPX KPI
```{r}
# Creating OPX KPI
KPI_OPX_num <- reactive({
  # testing for data quality, returning NA if bad
  if (datatest() == TRUE) {
    NA
  } else {
    exceldata() |>
      # most recent year
      filter(year == max(year)) |>
      # Calculating the value as a %
      transmute(
        KPI_OPX =
          sum(revenue_private_contributions, revenue_investments, assets_cash_in_bank_and_cash_equivalents, na.rm = TRUE) /
            sum(liability_short_term_grants_payable, liability_revocable_endowments, liability_long_term_grants_payable, na.rm = TRUE) * 100
      ) |>
      # Just the single number
      pull() |>
      first()
  }
})
```

@@

#### public/private proportion
```{r}
# Creating public/private proportion
KPI_proportion <- reactive({
  # testing for data quality, returning NA if bad
  if (datatest() == TRUE) {
    NA
  } else {
    exceldata() |>
      transmute(
        proportion = sum(revenue_private_contributions, revenue_membership_fees, na.rm = TRUE) / sum(revenue_public_grants, revenue_private_contributions, revenue_membership_fees, na.rm = TRUE) * 100
      ) |>
      pull() |>
      first()
  }
})
```

@@

## Function to read Excel file -------------------------------------------------
widen_excel <- function(file) {
  read_excel(file) |>
    # uniting Category and Name to deal with multiple 'Other'
    unite(
      entry, c(Category, Name),
      sep = "_"
    ) |>
    # deleting Sub-category and Firm
    select(-`Sub-category`, -Firm) |> 
    # pivoting to create 'tidy' variables
    pivot_wider(names_from = entry, values_from = Value) |>
    # cleaning names
    clean_names() 
}

@@

##### no_information
```{r}
## no_information plot ---------------------------------------------------------
# Function to print a 'no information' plot, if necessary
no_information <- function() {
  empty_plot("No information available")
}
```

##### bad_data
```{r}
## bad_data plot ---------------------------------------------------------------
# Function to print a 'bad data' plot, if necessary
bad_data <- function() {
  empty_plot("Something is wrong\nwith the Excel data.")
}
```

#### empty_plot
```{r}
## empty_plot-------------------------------------------------------------------
# Function to print an empty plot with a supplied message
empty_plot <- function(warning) {
  ggplot() +
    # text geom with a space to create the plot, but otherwise show nothing
    geom_text(aes(0, 0, label = " ")) +
    # as blank a slate as possible
    theme_void() +
    annotate(
      "text",
      label = warning,
      # Using the same palette as the other non-profit/loss plots
      x = 0, y = 0, size = 8, colour = met.brewer("Hokusai2", 1)
    )
}
```
